---
title: "Extract SST from NOAA High-resolution Blended Analysis and calculate DHW"
author: "Ana Palacio"
date: "September 9, 2019"
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script use the .nc files obtained from ftp://ftp.cdc.noaa.gov/Datasets/noaa.oisst.v2.highres/ (data from 1981 to 2014 in the "DHW/Daily_OISST_v2" folder) to extract the SST data for the locations of interest (coordinates = "DHW/Data/Coordinates.csv"), and then matches the maximum DHW during a year with the rate of change of coral coral measured in the location.

```{r}
rm(list = ls())  # Remove all objects from memory
setwd("DHW")# Move working directory inside the DHW folder
```

# 1. SST Data 

### Extract temperature data based on the coordinates

```{r libraries1, message=FALSE, results="hide"}
  library(raster)
  library(parallel)
```
 
```{r Extract_SST_data, cache=TRUE}

   # read all the .nc files 
    flist<-list.files('Daily_OISST_v2',pattern='*.nc',full.names=TRUE)
      
  # brick them 
    ras<-lapply(flist, stack)
    pts <- read.csv("Data/Coordinates.csv", stringsAsFactors = FALSE)
    extract.pts <- cbind(lon=pts$lon, lat=pts$lat)
    
  ## Extract data by coordinates 
    ext <- lapply(ras,raster::extract,extract.pts)
    #head(ext)
    SST <- unlist(ext, recursive = FALSE)
    #head(SST)
```


### Bind temperature data with coordinates and dates

```{r Bind_SST_GPS_Date}
  
# Create a date list 
    Dates<-seq.Date(as.Date("1981-09-01"), as.Date("2014-12-31"), by = "day")
    Date<-as.character(Dates)
    #head(Date)
      
# Bind GPS points and Dates
    Date.location <- merge(pts, Date)
    Temperature.data<-data.frame(Date.location, SST)
    Temperature.data$Date<-as.Date(Temperature.data$y)
    # head(Temperature.data)  
    # summary(Temperature.data)
```

# 2. DHW calculation

```{r libraries2, results="hide"}
  library(lubridate)
  library(dplyr)
  library(tidyverse)
  library(zoo)
  library(reshape2)
```

### Calculate MMM

For the climatology and maximum monthly mean (MMM) we selected the SST data from 1985-2012, to keep consistency with the CRW methodology

```{r Climatology}
  Temperature.data$Month<-month(Temperature.data$Date)
  Temperature.data$Year<-year(Temperature.data$Date)
  Temperature.data$Site<-as.factor(Temperature.data$Site) 
  
  # SELECT YEARS 1985-2012
    SST.1985_2012<-filter(Temperature.data, Year>1984)
    SST.1985_2012<-filter(SST.1985_2012, Year<2013)

  # Mean temperature per month in each location
  OI.MonthlyClimatology_1985_2012<-SST.1985_2012 %>%
    group_by(Site, Month) %>%
    summarize(Clima_M = mean(SST, na.rm = TRUE))
  #OI.MonthlyClimatology_1985_2012

  # Maximum Monthly Mean in each location
  MMM_OI<-OI.MonthlyClimatology_1985_2012 %>%
    group_by(Site) %>%
    summarize(MMM = max(Clima_M, na.rm = TRUE))
  # MMM_OI
```

### Calculate HotSpots

```{r HotSpots}

Temperature.data <- merge(Temperature.data, MMM_OI, on="Site") # Add the MMM to each location
Temperature.data <- Temperature.data[order(Temperature.data$Site, Temperature.data$Date),]

Temperature.data$HotSpot<-(Temperature.data$SST-(Temperature.data$MMM))
Temperature.data$HotSpot<-ifelse(Temperature.data$HotSpot>=0,
                                Temperature.data$HotSpot, 0)# REMOVE NEGATIVE ANOMALIES
Temperature.data$D_Stress<-ifelse(Temperature.data$HotSpot>=1,
                               Temperature.data$HotSpot, 0) # REMOVE HotSpots <1
Temperature.data$W_Stress<-(Temperature.data$D_Stress/7) # Transform daily to weekly values
```

### Calculate DHW

```{r DHW}
# Rolling sum of 84 days (12 weeks) Hotspots for each location
DHW_Daily<-Temperature.data %>% 
  group_by(Site) %>%
  mutate(NOAA_DHW_OI=rollapplyr(W_Stress,list(-(83:0)),sum,fill=NA, partial=FALSE)) %>%
  ungroup
#write.csv(DHW_Daily, "DHW_Daily.csv")
```

### Summarise max DHW and max SST by location and year

```{r}
# 1. Summary by Site and year (max values)
  DHW_Year_Summary<-DHW_Daily %>%
         group_by(Site, Year) %>%
         dplyr::summarise(maxDHW = max(NOAA_DHW_OI),
                          maxSST = max(SST))

# 2. Dulplicate places that belong to the same OISST pixel in ordet to match them with changes in coral cover  
    # Gorgona_Island = La_Azufrada
    La_Azufrada<-filter(DHW_Year_Summary, Site=="Gorgona_Island")
    La_Azufrada$Site<-"La_Azufrada"
    
    # Wolf_North_Corals = Wolf_South_Corals
    Wolf_South_Corals<-filter(DHW_Year_Summary, Site=="Wolf_North_Corals")
    Wolf_South_Corals$Site<-"Wolf_South_Corals"
    
    # Roca_Espejo_1 = Roca_Espejo_2
    Roca_Espejo_2<-filter(DHW_Year_Summary, Site=="Roca_Espejo_1")
    Roca_Espejo_2$Site<-"Roca_Espejo_2"
    
    # Bahia_Banderas = Tenacatita
    Tenacatita<-filter(DHW_Year_Summary, Site=="Bahia_Banderas")
    Tenacatita$Site<-"Tenacatita"
    
    DHW_Year_Summary<-rbind(DHW_Year_Summary, La_Azufrada,
                            Wolf_South_Corals, Roca_Espejo_2, 
                            Tenacatita)    
    
    write.csv(DHW_Year_Summary, "Output/maxDHW.csv")
    
#3. Wide format data. Not required for further analysis
    #DHW_Year<-DHW_Year_Summary[, -4]
    #DHW_All_wide <-spread(DHW_Year, Year, maxDHW)
    #DHW_All_wideb <-spread(DHW_Year, Site, maxDHW)
    # write.csv(DHW_All_wide, "Output/maxDWH_Wide.csv")
```


### Merge maxDHW data with Rate of Coral cover data

```{r Summaries}

# 1. Add "Start_year" variable
    DHW_Year_Summary$Start_year<-DHW_Year_Summary$Year

# 2. Read the rate of cover file and merge
    CC_Change<-read.csv("Data/Rate_of_change_cover.csv", header = TRUE)
    RoC_DHW<-left_join(CC_Change, DHW_Year_Summary, by=c("Start_year", "Site"))
    write.csv(RoC_DHW, "Output/RoC_DHW.csv")
    setwd("..") # move back to main project folder to continue with coral cover analysis
```
